\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-04-21 honors_thesis/journal_paper/paper.tex
Started moving the results to be in one group at the end

The balance of this paper is organized as follows. \sec{pso} describes the
particle swarm optimization algorithm.  \sec{sepso} describes how speculative
evaluation can be done in parallel PSO to perform two iterations at once, then
presents a comparison of the speculative algorithm with the original PSO.  In
\sec{relax}, we discuss various methods of improving the performance of
speculative evaluation in PSO, all of which break the requirement of strictly
reproducing the behavior of the original algorithm.  In \sec{conclusion} and
\sec{future} we conclude and discuss future work.

@@

\subsection{Comparison with Standard PSO}
\label{sec:results}

Lastly we look at a function that does not show any kind of improvement with
this method.  The function Rastrigin is defined as $f(\Vec{x}) = \sum_{i=1}^D
\left(x_i^2 - 10\cos\left(2\pi x_i\right) + 10\right)$.  It has been shown that
with Rastrigin, the more particles there are in the swarm, the lower function
value it finds, up to at least 4000
particles~\cite{mcnabb-2009-large-particle-swarms}.  Smaller swarms get caught
in local optima.  Because our speculative algorithm has a significantly smaller
swarm size, it gets stuck at higher values while the larger swarms performing
regular PSO continue to improve the best value found.  \fig{rastrigin} shows
the results graphically.

@@

While the idea of speculative evaluation in particle swarm optimization is
intriguing, the results of the previous section show that in many cases it
requires too many extra evaluations to produce competitive results.  On most
functions, the larger swarm size that can be used in regular PSO leads to
better performance than is found by trying to do speculative evaluation.
However, if we keep the idea of speculative evaluation while relaxing the
requirement of exactly reproducing the behavior of the original PSO algorithm,
we see some impressive results.

@@

@@

>>>>

The balance of this paper is organized as follows. \sec{pso} describes the
particle swarm optimization algorithm.  \sec{sepso} describes how speculative
evaluation can be done in parallel PSO to perform two iterations at once.  In
\sec{relax}, we discuss various methods of improving the performance of
speculative evaluation in PSO, all of which break the requirement of strictly
reproducing the behavior of the original algorithm.  \sec{results} presents the
results of using our methods in comparison with standard PSO.  In
\sec{conclusion} and \sec{future} we conclude and discuss future work.

@@

@@

While the idea of speculative evaluation in particle swarm optimization is
intriguing, the results in \sec{results} show that in many cases it requires
too many extra evaluations to produce competitive results.  On most functions,
the larger swarm size that can be used in regular PSO leads to better
performance than is found by trying to do speculative evaluation.  However, if
we keep the idea of speculative evaluation while relaxing the requirement of
exactly reproducing the behavior of the original PSO algorithm, we see some
impressive results.

@@

\section{Results}
\label{sec:results}

We present results for five common benchmark problems: Rastrigin, Sphere,
Schwefel 2.21, Griewank, and Bohachevsky.  Rastrigin is defined as:
\[f(\Vec{x}) = \sum_{i=1}^D\left(z_i^2 - 10\cos\left(2\pi z_i\right) +
10\right), z=x-c\]
initialized in $[-5.12,5.12]^D$.
Sphere is defined as:
\[f(\Vec{x}) = \sum_{i=1}^D z_i^2, z=x-c\]
initialized in $[-50,50]^D$.
Schwefel 2.21 is defined as:
\[f(\Vec{x}) = \max_i{|z_i|, 1 \leq i \leq D}, z=x-c\]
initialized in $[-500,500]^D$.
Griewank is defined as:
\[f(\Vec{x}) = \frac{1}{4000}\sum_{i=1}^D z_i^2 - \Pi_{i=1}^D
\cos\left(\frac{z_i}{\sqrt{i}} \right) + 1, z=x-c\]
initialized in $[-600,600]^D$
And Bohachevsky is defined as:
\[f(\Vec{x}) = \sum_{i=1}^D(x_i^2+2x_{i+1}^2-.3\cos(3\pi x_i)-.4\cos(4\pi
x_{i+1})+.7)\]
initialized in $[-15,15]^D$.

The $c$ in the function definitions is a shifted center, in order to avoid
origin-seeking bias in the PSO algorithm and its
variants~\cite{monson-2005-origin-seeking-bias}.  We move the center of the
shifted functions (all except Bohachevsky) to be halfway between the center and
the boundary of the initialization region.  For example, if the initialization
region is $[-50,50]^D$, the center is $(25)^D$.

Lastly we look at a function that does not show any kind of improvement with
this method.  The function Rastrigin is defined as 
It has been shown that
with Rastrigin, the more particles there are in the swarm, the lower function
value it finds, up to at least 4000
particles~\cite{mcnabb-2009-large-particle-swarms}.  Smaller swarms get caught
in local optima.  Because our speculative algorithm has a significantly smaller
swarm size, it gets stuck at higher values while the larger swarms performing
regular PSO continue to improve the best value found.  \fig{rastrigin} shows
the results graphically.
\subsection{Comparison with Standard PSO}

\subsection{Pick Best}

\subsection{Pruning}

@@

\subsection{Many Iters}
